---
title: "Homework-1"
author: "Christopher LaBorde"
date: '2022-04-03'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

Homework 1
PSTAT 131/231
Machine Learning Main Ideas

Please answer the following questions. Be sure that your solutions are clearly marked and that your document is neatly formatted.

You don’t have to rephrase everything in your own words, but if you quote directly, you should cite whatever materials you use (this can be as simple as “from the lecture/page # of book”).
Question 1:

Define supervised and unsupervised learning. What are the difference(s) between them?

**Supervised learning uses observed outcomes as a "supervisor" or validator of the effectiveness of a model, comparing the predictions made by the model to actual values, while unsupervised learning looks at the data and the outcomes in order to discern or deduce the rules that dictate the relationships between predictor variables and the outcomes**

Question 2:

Explain the difference between a regression model and a classification model, specifically in the context of machine learning. 

**A regression model is a mathematical expression which attempts to explain the relationship between predictor variables and a response variable. With regards to machine learning, regression models are suited when the response variable is quantitative and continuous. A classification model is a model which has a range of discretely-valued outputs. With regards to machine learning, classification models are good for models which have qualitative response variables.**

Question 3:

Name two commonly used metrics for regression ML problems. Name two commonly used metrics for classification ML problems.

**2 Commonly used metrics for regression ML problems are Training Mean Squared Error (MSE) and Test MSE.2 commonly used metrics used for classification ML problems are Training error rate and the test error rate.**


Question 4:

As discussed, statistical models can be used for different purposes. These purposes can generally be classified into the following three categories. Provide a brief description of each.

    Descriptive models:
  **visually emphasises a trend in data - i.e. a line on a scatterplot** 

    Inferential models: 
  **States relationship between response and predictors, attempts to identify which features are significant and aims to test theories**

    Predictive models: 
  **attempts to predict response with minimum reducible error**
    
*slide 7, PSTAT 131 lecture day 2*


Question 5:

Predictive models are frequently used in machine learning, and they can usually be described as either mechanistic or empirically-driven. Answer the following questions.

    Define mechanistic. Define empirically-driven. How do these model types differ? How are they similar?
    
**A mechanistic predictive model is also called a parametric model, and assumes that the relationship between predictor and response variables can be expressed by parameters (coefficients) within the model. A model described as empirically driven is one that is created by observing the data and deducing the relationships, rather than finding relationships by theory. Empirically-driven models are more prone to overfitting than mechanistic models. They are similar in that they both attempt to describe the relationship between predictors and response**

    In general, is a mechanistic or empirically-driven model easier to understand? Explain your choice.

**In general, a mechanistic model is easier to understand than an empirically driven one. This is because the parameters or coefficients can be laid out explicitly and expressed in a mathematical expression that is readable and relatively intuitive, represented by weights or relative effects of each predictor variable on the value of the response.**

    Describe how the bias-variance tradeoff is related to the use of mechanistic or empirically-driven models.
    
**empirically-driven models are more likely to overfit. This means that the variance of the model will be higher, but the bias of the model will be less. On the other hand, a mechanistic model, when ideally fitted will have a higher bias but a lower variance.**



Question 6:

A political candidate’s campaign has collected some detailed voter history data from their constituents. The campaign is interested in two questions:

    Given a voter’s profile/data, how likely is it that they will vote in favor of the candidate?
  **I would classify this question as predictive. Relating to the definition given in question 4, here we are trying to predict the response with minimal error**

    How would a voter’s likelihood of support for the candidate change if they had personal contact with the candidate?
  **I would classify this question as inferential, as here we are trying to determine which predictors are influential, in this case the predictor under scrutiny being whether or not a voter had personal contact with a candidate.**

Classify each question as either predictive or inferential. Explain your reasoning for each.


Exploratory Data Analysis

This section will ask you to complete several exercises. For this homework assignment, we’ll be working with the mpg data set that is loaded when you load the tidyverse. Make sure you load the tidyverse and any other packages you need.

```{r}
library(tidyverse)
library(ggplot2)
library(corrplot)
```

Exploratory data analysis (or EDA) is not based on a specific set of rules or formulas. It is more of a state of curiosity about data. It’s an iterative process of:

    generating questions about data
    visualize and transform your data as necessary to get answers
    use what you learned to generate more questions

A couple questions are always useful when you start out. These are “what variation occurs within the variables,” and “what covariation occurs between the variables.”

You should use the tidyverse and ggplot2 for these exercises.
Exercise 1:

We are interested in highway miles per gallon, or the hwy variable. Create a histogram of this variable. Describe what you see/learn.

```{r}
mpg

ggplot(mpg, aes(x = hwy)) + geom_histogram()
```

The distribution for highway miles per gallon "hwy" in the dataset looks to be bi-modal (has two peaks), with the largest amount of cars having a highway mpg of 26, and the second peak being centered at 17 mpg. The highest highway mpg observed in the dataset is 44 mpg, the lowest is 13 mpg. 


Exercise 2:

Create a scatterplot. Put hwy on the x-axis and cty on the y-axis. Describe what you notice. Is there a relationship between hwy and cty? What does this mean?

```{r}
ggplot(mpg, aes(y = cty,x = hwy)) + geom_point()
```
The plot shows an obvious increasing linear relationship between the variables hwy and cty. I am assuming cty stands for miles per gallon when driven in the city. This means that a vehicle with a higher hwy mpg will likely have a higher cty mpg than other cars as well.


Exercise 3:

Make a bar plot of manufacturer. Flip it so that the manufacturers are on the y-axis. Order the bars by height. Which manufacturer produced the most cars? Which produced the least?


```{r}
library(data.table)
data <- mpg
df <- data.table(mpg)

df[,manufacturer := as.factor(manufacturer)]


ggplot(mpg,aes(y = fct_infreq(manufacturer))) + 
	geom_bar(stat = 'count') + ylab('manufacturer')

# reordering process learned at: https://www.roelpeters.be/reorder-ggplot2-bar-chart-by-count/

```

Dodge produced the most cars, while Lincoln produced the least.


Exercise 4:
Make a box plot of hwy, grouped by cyl. Do you see a pattern? If so, what?

```{r}
boxplot(mpg$hwy ~ mpg$cyl, xlab = "No. Cylinders", ylab = "highway mpg")

```
It seems that as the number of cylinders increases, the average, maximum,and minimum values for highway mpg generally decrease.



Exercise 5:

Use the corrplot package to make a lower triangle correlation matrix of the mpg dataset. (Hint: You can find information on the package here.)

Which variables are positively or negatively correlated with which others? Do these relationships make sense to you? Are there any that surprise you?


```{r}
str(mpg)
corr <- cor(mpg[c('displ', 'year', 'cyl', 'cty', 'hwy')])

corrplot(corr, type = 'lower')
```
For exercise 5, I decided to include only the variables that were of a numeric type. Following are the pairs of positive and negative correlations, with perceived reasons for correlation.

Pairs:
+ correlated:
(hwy, cty) - a vehicles city mpg will be higher if its hwy mpg is higher
(cyl, displ) - as the number of cylinders increases, so does the displacement (generally)

- coorrelated:
(cty, displ) - the larger an engine's displacement, the lower mpg (generally)
(hwy, displ) - the larger an engine's displacement, the lower mpg (generally)
(cty, cyl) - the more cylinders in an engine, the lower its mpg (generally)
(hwy, cyl) - the more cylinders in an engine, the lower its mpg (generally)




231 Students Only:
Exercise 6:

Recreate the following graphic, as closely as you can. Hint: Use the ggthemes package.




```{r}
ggplot(mpg, aes(x = hwy, y = class)) + geom_boxplot() + geom_jitter(width = 0) + xlab('Highway MPG') + ylab('Vehicle Class')

```


Exercise 7

Recreate the following graphic.

```{r}
ggplot(mpg, aes(x = class, y = hwy, fill = drv)) + geom_boxplot()

```

Exercise 8

Recreate the following graphic.

```{r}
ggplot(mpg, aes(x = displ, y = hwy, color = drv)) + geom_point() + geom_smooth(linetype = 'dashed', se = FALSE)

```
